Fast and Flexible Binary Instrumentation for the Kernel

### 2013-11-25 VEE-14/paper.tex
Working on masters paper; writing about DBT optimizations in Granary.

\subsection{Optimizations}
\paragraph{Ahead-of-time Translation} 
Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s. Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e. when the conditional jump is not taken).

@@

The method used to pad basic blocks so as to align specific instructions must be transparent (i.e. not change the execution behavior). The typical approach to alignment is to introduce \texttt{nop} instructions, or \texttt{jmp} instructions which jump over the padding. Adding extra instructions, however, introduces added latency to the fetch/decode/execute cycle. Therefore, when possible, Granary pads instructions using otherwise ignored instruction prefixes. For example, Granary aligned patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}).

@@

\paragraph{Trace Allocation}

>>>>

\subsection{Optimizations}
\paragraph{Ahead-of-time Translation} \label{sec:aot}
Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s. Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e. when the conditional jump is not taken).

@@

%Padding basic blocks so as to align specific instructions must not change the execution behavior.

One approach to change an instruction's alignment is to pad the basic block with additional \texttt{nop} instructions. An alternative when more than a few bytes of padding are needed is to introduce a single control-flow instruction (\texttt{jmp}) that jumps around the padded area. However, adding extra instructions introduces latency into the fetch/decode/execute cycle. Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment. For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}. Finally, most basic blocks are small (i.e. fit in one cache line), therefore, aligning basic blocks on a cache line boundary naturally avoids the potential for a patchable instruction to straddle two cache lines.

@@

\paragraph{Trace Allocation}
Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}. Typical approaches to tracing require a runtime profiling infrastructure (e.g. recording basic block execution counts). Granary attempts to reap the benefits of tracing without the runtime profiling overhead by opportunistically forming traces using ahead-of-time translation (\Cref{sec:aot}), and by arranging for related basic blocks to be allocated sequentially in memory.

Two basic blocks $B_1$ and $B_2$ are ``related" ($B_1 \to B_2$) if there is a control flow instruction from $B_1$ to $B_2$. Grnary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions. The base case of this induction are kernel entrypoints. For example, each system call is an entrypoint into the kernel, and is associated with a private memory pool. When a system call is executed for the first time, the basic blocks needed to fulfill the system call's computation are allocated from that system call's memory pool.

To avoid translation conflicts (e.g. on utility functions being simultaneously executed by distinct system calls), and to improve the predictability and reproducability of a given code cache layout, Granary can delay the takeover of kernel entrypoints. That is, Granary can take over one entrypoint at a time, and wait for a specified period of time before taking over the next entrypoint. The effect of delaying the takeover process is that it increases the chances that a given entrypoint (e.g. a blocking system call) will execute to completion, and thus have all of its basic blocks  arranged sequentially in memory.

%, and is associated with a private memory pool from which to allocate memory for basic blocks. The net result is that most of the basic blocks eecuted in service of a single system call are allocated from that system call's memory pool.

%The intuition for this approach is that kernel code is not very ``loopy", and so

%Granary services memory allocation requests for basic block instructions from memory pools. By default, Granary has as many memory pools as there are CPUs. However, when trace allocation is enabled

%Tracing is also beneficial for

%However, tracing of this sort requires mechnisms for evicting stale basic blocks,

%Constructing traces of ``hot'' 
%JIT generally results in better performance .

%By default, Granary allocates the memory needed to hold the instructions of a basic block from CPU-private memory pools. Each memory pool serves basic block allocation requests out of fixed-size memory blocks, ranging from 4K to 24K in size.
